---
title: "Social Data Science Project Analysis"
authors: "Andreas Opedal & Charlotte Out"
date: "4/16/2020"
output: html_document
---

```{r}
# Load packages
library(dplyr)
library(syuzhet)
library(WDI)
library(tools)
library(sentimentSetsR)
```

Before the sentiment and correlation analysis can be done a couple of pre-processing steps are needed.

```{r}
# Load and clean vegan tweet data

vegan_data <- read.csv("vegan_tweets_final.csv")
vegan_data <- select(vegan_data, -X)
vegan_data$location <- unlist(lapply(vegan_data$location, as.character))
vegan_data$country <- unlist(lapply(vegan_data$country, as.character))
vegan_data$text <- unlist(lapply(vegan_data$text, as.character))
vegan_data$lang <- unlist(lapply(vegan_data$lang, as.character))

```


```{r}
# Change country strings to title case
vegan_data$country <- toTitleCase(vegan_data$country)

# Take a look at distribution of observations over countries

vegan_data %>% 
  group_by(country) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) -> vegan_country_groupings

# View(vegan_country_groupings)

```

***
Approach: Restrict ourselves to tweets in English


```{r}
# Check distribution of tweets in english

vegan_data %>%
  filter(lang == "en") -> eng_tweets

eng_tweets %>% 
  group_by(country) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) -> eng_tweets_by_country

View(eng_tweets_by_country)

# Set threshold to 100 observations

eng_tweets_by_country %>%
  filter(count >= 100) -> eng_tweets_by_country_filtered

print(paste0("Observe that there are ", nrow(eng_tweets_by_country_filtered), " countries left"))

eng_tweets %>%
  filter(country %in% eng_tweets_by_country_filtered$country) -> eng_tweets

```


```{r}

## Note that the Syuzhet scores are used for benchmarking only, not for analysis

# get sentiments of tweets from Syuzhet and aggregate by country on average

eng_tweets$Syuzhet <- get_sentiment(eng_tweets$text)

eng_tweets %>%
  group_by(country) %>%
  summarize(vegsentiment_syuzhet = mean(Syuzhet)) -> eng_tweets_avg


```


```{r}

# get sentiments of tweets from VADER and aggregate by country on average

eng_tweets$VADER <- NA
for (i in seq(1, nrow(eng_tweets)))
{
  eng_tweets$VADER[i] <- getVaderRuleBasedSentiment(eng_tweets$text[i], compound=T)
}

eng_tweets %>%
  group_by(country) %>%
  summarize(vegsentiment_vader = mean(VADER)) -> eng_tweets_avg_2

#View(eng_tweets_avg_2)

# Join with Syuzhet
inner_join(eng_tweets_avg, eng_tweets_avg_2) -> eng_tweets_avg

```


```{r}

eng_tweets_avg <- read.csv("vegan_sentiment_averages.csv")

```

Now load and pre-process GDP/capita, Future Orientation Index, Global Innovation Index, Human Development Index and Climate Change Performance Index. 

Start with GDP/capita (corrected by PPP) which can be retrieved from the World Bank. Want to find the most recent year's data. (Note HDI index is also in the World Bank's database but had only NA values for recent years)

```{r}
# Search for gdp/capita
gdpsearch <- WDIsearch(string = "gdp per capita")

# Select "GDP per capita, PPP (constant 2011 international $)" ("NY.GDP.PCAP.PP.KD")
(gdp <- gdpsearch[[7]])

# Search for HDI
# (hdisearch <- WDIsearch(string = "human development"))

# Select "Human development index (HDI)", ("UNDP.HDI.XD")
# (hdi <- hdisearch[[1]])

# Now download data 
WDIdf <- WDI(indicator = c(gdp), start = 2018, end = 2018)

# Filter on countries in our data set

WDIdf %>%
  filter(country %in% eng_tweets_avg$country) -> WDIdf
#View(WDIdf)

# Note: assert that no countries are missing
(nrow(WDIdf) == nrow(eng_tweets_avg))

# Join on country
inner_join(WDIdf, eng_tweets_avg) %>%
  select(country, vegsentiment_syuzhet, vegsentiment_vader, NY.GDP.PCAP.PP.KD) %>%
  rename(GDP_capita = NY.GDP.PCAP.PP.KD) -> agg_data

# Compute Pearson correlation
(obs_gdp <- cor(agg_data$vegsentiment_vader, agg_data$GDP_capita))

```

Now load HDI data

```{r}

hdi <- read.csv("Human Development Index (HDI).csv", sep = ";")
#View(hdi)

# Select only most recent data and filter on countries in our dataset
hdi %>%
  select(Country, X2018) %>%
  filter(Country %in% eng_tweets_avg$country) %>%
  rename(HDI = X2018) -> hdi

# Assert that no country is missing
(nrow(hdi) == nrow(eng_tweets_avg))

# Join on country
inner_join(agg_data, hdi, by = c("country" = "Country")) -> agg_data

# Fix problems with data type
agg_data$HDI <- unlist(lapply(agg_data$HDI, as.character))
agg_data$HDI <- unlist(lapply(agg_data$HDI, as.double))

# Compute Pearson correlation
(obs_hdi <- cor(agg_data$vegsentiment_vader, agg_data$HDI))

```

Load GII data

```{r}
gii <- read.csv("Global Innovation Index.csv")

#View(gii)

gii %>%
  select(Economy, Score) %>%
  rename(country = Economy, GII = Score) -> gii

gii$country <- unlist(lapply(gii$country, as.character))

# Modify country names
gii[gii$country == "Netherlands (the)",]$country <- "Netherlands"
gii[gii$country == "United Kingdom (the)",]$country <- "United Kingdom"
gii[gii$country == "United States of America (the)",]$country <- "United States"

# Filter out countries
gii %>% 
  filter(country %in% agg_data$country) -> gii

# Assert that no country is missing
(nrow(gii) == nrow(eng_tweets_avg))

# Join
inner_join(agg_data, gii) -> agg_data

# Compute Pearson correlation
(obs_gii <- cor(agg_data$vegsentiment_vader, agg_data$GII))

```

Climate Change Performance Index

```{r}
CCPI <- read.csv("CCPI.csv", sep = ";")
#View(CCPI)

CCPI$country <- unlist(lapply(CCPI$country, as.character))

# Filter out countries
CCPI %>%
  filter(country %in% agg_data$country) -> CCPI

# Join 
inner_join(agg_data, CCPI) -> agg_data

# Compute Pearson correlation
(obs_ccpi <- cor(agg_data$vegsentiment_vader, agg_data$CCPI))

```

Future Orientation Index

```{r}
# Load and pre-process
GoogleDF <-read.csv("geoMap.csv", skip=2)
names(GoogleDF) <- c("country", "G2018", "G2020")
GoogleDF$G2018 <- as.numeric(sub("%", "", GoogleDF$G2018))
GoogleDF$G2020 <- as.numeric(sub("%", "", GoogleDF$G2020))

# Compute FOI
GoogleDF$FOI <- GoogleDF$G2020/GoogleDF$G2018

# Filter on countries in data set and remove unnecesarry columns
GoogleDF %>%
  filter(country %in% agg_data$country) %>%
  select(country, FOI) -> GoogleDF

# Assert no country is missing
(nrow(GoogleDF) == nrow(eng_tweets_avg))

# Join
inner_join(agg_data, GoogleDF) -> agg_data

# Compute Pearson correlation
(obs_foi <- cor(agg_data$vegsentiment_vader, agg_data$FOI))

```


Now perform permutation tests to assess how plausible it is that our observations could have been produced by chance under null hypothesis of no correlation.

Alternative hypothesis is as we recall a positive correlation for all factors. 


```{r}
# Permutation test for Pearson correlation

nsim <- 1000
gdpcor <- c()
hdicor <- c()
giicor <- c()
ccpicor <- c()
foicor <- c()
for (i in seq(nsim)) {
  shufdata <- agg_data[sample(nrow(agg_data)),]
  gdpcor <- append(gdpcor, cor(shufdata$vegsentiment_vader, agg_data$GDP_capita))
  hdicor <- append(hdicor, cor(shufdata$vegsentiment_vader, agg_data$HDI))
  giicor <- append(giicor, cor(shufdata$vegsentiment_vader, agg_data$GII))
  ccpicor <- append(ccpicor, cor(shufdata$vegsentiment_vader, agg_data$CCPI))
  foicor <- append(foicor, cor(shufdata$vegsentiment_vader, agg_data$FOI))
}

```

```{r}
#Creating the histograms and calculating the p-values for the Pearson correlation coefficients 


# GDP/capita
hist(gdpcor,
 xlim=range(c(gdpcor, obs_gdp)), freq = F, 
 main = "Vegan sentiment and GDP/capita", xlab = "Pearson correlation coefficient")
abline(v=obs_gdp, col="red")

p_gdp <- (1 + sum(gdpcor >= obs_gdp)) / (length(gdpcor) + 1)

print(paste0("The correlation coefficient between vegan sentiment and GDP/capita is: ", round(obs_gdp, digits = 4)))
print(paste0("The p-value of the correlation between vegan sentiment and GDP/capita is: ", round(p_gdp, digits = 4)))

# HDI
hist(hdicor,
 xlim=range(c(hdicor, obs_hdi)), freq = F, 
 main = "Vegan sentiment and HDI", xlab = "Pearson correlation coefficient")
abline(v=obs_hdi, col="red")

p_hdi <- (1 + sum(hdicor >= obs_hdi)) / (length(hdicor) + 1)

print(paste0("The correlation coefficient between vegan sentiment and HDI is: ", round(obs_hdi, digits = 4)))
print(paste0("The p-value of the correlation between vegan sentiment and HDI is: ", round(p_hdi, digits = 4)))

# GII
hist(giicor,
 xlim=range(c(giicor, obs_gii)), freq = F, 
 main = "Vegan sentiment and GII", xlab = "Pearson correlation coefficient")
abline(v=obs_gii, col="red")

p_gii <- (1 + sum(giicor >= obs_gii)) / (length(giicor) + 1)

print(paste0("The correlation coefficient between vegan sentiment and GII is: ", round(obs_gii, digits = 4)))
print(paste0("The p-value of the correlation between vegan sentiment and GII is: ", round(p_gii, digits = 4)))

# CCPI
hist(ccpicor,
 xlim=range(c(ccpicor, obs_ccpi)), freq = F, 
 main = "Vegan sentiment and Climate Change Performance", xlab = "Pearson correlation coefficient")
abline(v=obs_ccpi, col="red")

p_ccpi <- (1 + sum(ccpicor >= obs_ccpi)) / (length(ccpicor) + 1)

print(paste0("The correlation coefficient between vegan sentiment and CCPI is: ", round(obs_ccpi, digits = 4)))
print(paste0("The p-value of the correlation between vegan sentiment and CCPI is: ", round(p_ccpi, digits = 4)))

# FOI
hist(foicor,
 xlim=range(c(foicor, obs_foi)), freq = F, 
 main = "Vegan sentiment and Future Orientation Index", xlab = "Pearson correlation coefficient")
abline(v=obs_foi, col="red")

p_foi <- (1 + sum(foicor >= obs_foi)) / (length(foicor) + 1)

print(paste0("The correlation coefficient between vegan sentiment and FOI is: ", round(obs_foi, digits = 4)))
print(paste0("The p-value of the correlation between vegan sentiment and FOI is: ", round(p_foi, digits = 4)))
```

```{r}
#Calculating spearman correlation coefficients 

obs_GDP_S <- cor(agg_data$vegsentiment_vader, agg_data$GDP_capita, method = "spearman")
obs_HDI_S <- cor(agg_data$vegsentiment_vader, agg_data$HDI, method = "spearman")
obs_GII_S <- cor(agg_data$vegsentiment_vader, agg_data$GII, method = "spearman")
obs_CCPI_S <- cor(agg_data$vegsentiment_vader, agg_data$CCPI, method = "spearman")
obs_FOI_S <-  cor(agg_data$vegsentiment_vader, agg_data$FOI, method = "spearman")

```

```{r}
# Permutation test using Spearman correlation coefficient 

nsim <- 1000
gdpcorS <- c()
hdicorS <- c()
giicorS <- c()
ccpicorS <- c()
foicorS <- c()
for (i in seq(nsim)) {
  shufdata <- agg_data[sample(nrow(agg_data)),]
  gdpcorS <- append(gdpcorS, cor(shufdata$vegsentiment_vader, agg_data$GDP_capita, method = 'spearman'))
  hdicorS <- append(hdicorS, cor(shufdata$vegsentiment_vader, agg_data$HDI, method = 'spearman'))
  giicorS <- append(giicorS, cor(shufdata$vegsentiment_vader, agg_data$GII, method = 'spearman'))
  ccpicorS <- append(ccpicorS, cor(shufdata$vegsentiment_vader, agg_data$CCPI, method = 'spearman'))
  foicorS <- append(foicorS, cor(shufdata$vegsentiment_vader, agg_data$FOI, method = 'spearman'))
}


```



```{r}
#creating the histrograms and calculating the p-values for Spearman correlation
#coefficient 


# GDP/capita
hist(gdpcorS,
 xlim=range(c(gdpcorS, obs_GDP_S)), freq = F, 
 main = "Vegan sentiment and GDP/capita", xlab = "Spearman correlation coefficient")
abline(v=obs_GDP_S, col="red")

p_gdpS <- (1 + sum(gdpcorS >= obs_GDP_S)) / (length(gdpcorS) + 1)

print(paste0("The spearman correlation coefficient between vegan sentiment and GDP/capita is: ", round(obs_GDP_S, digits = 4)))
print(paste0("The p-value of the correlation between vegan sentiment and GDP/capita is: ", round(p_gdpS, digits = 4)))

# HDI
hist(hdicorS,
 xlim=range(c(hdicorS, obs_HDI_S)), freq = F, 
 main = "Vegan sentiment and HDI", xlab = "Spearman correlation coefficient")
abline(v=obs_HDI_S, col="red")

p_hdiS <- (1 + sum(hdicorS >= obs_HDI_S)) / (length(hdicorS) + 1)

print(paste0("The spearman correlation coefficient between vegan sentiment and HDI is: ", round(obs_HDI_S, digits = 4)))
print(paste0("The p-value of the correlation between vegan sentiment and HDI is: ", round(p_hdiS, digits = 4)))

# GII
hist(giicorS,
 xlim=range(c(giicorS, obs_GII_S)), freq = F, 
 main = "Vegan sentiment and GII", xlab = "Spearman correlation coefficient")
abline(v=obs_GII_S, col="red")

p_giiS <- (1 + sum(giicorS >= obs_GII_S)) / (length(giicorS) + 1)

print(paste0("The Spearman correlation coefficient between vegan sentiment and GII is: ", round(obs_GII_S, digits = 4)))
print(paste0("The p-value of the correlation between vegan sentiment and GII is: ", round(p_giiS, digits = 4)))

# CCPI
hist(ccpicorS,
 xlim=range(c(ccpicorS, obs_CCPI_S)), freq = F, 
 main = "Vegan sentiment and Climate Change Performance", xlab = "Spearman correlation coefficient")
abline(v=obs_CCPI_S, col="red")

p_ccpiS <- (1 + sum(ccpicorS >= obs_CCPI_S)) / (length(ccpicorS) + 1)

print(paste0("The Spearman correlation coefficient between vegan sentiment and CCPI is: ", round(obs_CCPI_S, digits = 4)))
print(paste0("The p-value of the correlation between vegan sentiment and CCPI is: ", round(p_ccpiS, digits = 4)))

# FOI
hist(foicorS,
 xlim=range(c(foicorS, obs_FOI_S)), freq = F, 
 main = "Vegan sentiment and Future Orientation Index", xlab = "Spearman correlation coefficient")
abline(v=obs_FOI_S, col="red")

p_foiS <- (1 + sum(foicorS >= obs_FOI_S)) / (length(foicorS) + 1)

print(paste0("The Spearman correlation coefficient between vegan sentiment and FOI is: ", round(obs_FOI_S, digits = 4)))
print(paste0("The p-value of the correlation between vegan sentiment and FOI is: ", round(p_foiS, digits = 4)))


```


```{r}
# Check Pearson correlations with median sentiment

eng_tweets %>%
  group_by(country) %>%
  summarize(vegsentiment_vader_median = median(VADER)) -> eng_tweets_med

inner_join(agg_data, eng_tweets_med) -> agg_data

(cor(agg_data$vegsentiment_vader_median, agg_data$GDP_capita))
(cor(agg_data$vegsentiment_vader_median, agg_data$HDI))
(cor(agg_data$vegsentiment_vader_median, agg_data$GII))
(cor(agg_data$vegsentiment_vader_median, agg_data$CCPI))
(cor(agg_data$vegsentiment_vader_median, agg_data$FOI))

```




